---
title: "Homework 2"
author: Wuraola Olawole
output: github_document
---



```{r initial setup, message = FALSE}

    library(tidyverse)
      library(readxl)
```

# Problem 1

Set a path

```{r set_path}

  path_2_data = "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"

```

Read dataset, Mr. Trashwheel, set parameters, and modify sports_balls

```{r trash_df}

trashwheel_df =

    read_excel(
                path = path_2_data,
                  sheet = "Mr. Trash Wheel",
                    range = cell_cols("A:N"),
                      col_types = NULL,
                        skip = 0) %>%
        janitor::clean_names() %>%
          drop_na(dumpster) %>%
            mutate(
                  sports_balls = round(sports_balls),
                  sports_balls = as.integer(sports_balls)
  )
  
```
Read data on Precipitation in 2017 and in 2018 and add variable "year"

```{r}

path_2_data = "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"

  precipi_17_df =
  
  read_excel(
              path = path_2_data,
                sheet = "2017 Precipitation",
                  col_types = NULL,
                    skip = 1) %>%
    janitor::clean_names() %>%
      drop_na(month) %>%
        mutate(
                year = 2017) %>%
  relocate(year)

```
```{r}

precipi_18_df =
  
  read_excel(
                path = path_2_data,
                  sheet = "2018 Precipitation",
                    col_types = NULL,
                      skip = 1) %>%
  janitor::clean_names() %>%
    drop_na(month) %>%
      mutate(
              year = 2018) %>%
  relocate(year)

```
Combine precipitation dataset and Change month to Character variable

```{r month_df}

month_df =
    tibble(
            month = 1:12,
            month_name = month.name
    )
```
```{r Precip_17_18}

precipi_df = 
            bind_rows(precipi_17_df, precipi_18_df)

precipi_df =
            left_join(precipi_df, month_df, by = "month")

```
  Mr. TrashWheel data set contained data about trash that was collected in a dumpster (by Mr. TrashWheel) as it entered the inner harbor in Baltimore Maryland. The data set contained variables year and date of trash collection as well as the weights in tons and volume of the trash collected. It also provided information about the categories of trash collected in the dumpster. In this data set (final outcome), there were `r nrow (trashwheel_df)`. There were also additional sheets that provided information about precipitation; the month and total amount of precipitation in 2017 and 2018.

The total precipitation observed in 2018 was `r precipi_18_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches and the median number of sports balls in the dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.


# problem 2

Read the NYC~ csv data, retain select variables,convert 'entry' variable from character to logical variable using the 'recode' function.

```{r NYC_df, message = FALSE}

path_3_data = "./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
  
  NYC_df = 
    
    read_csv(
             path_3_data) %>% 
              janitor::clean_names() %>% 
                select(2:18, 20, 23) %>%
  mutate(
    entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) %>%
    mutate_at(vars(route8:route11), as.character)             # convert dbl variables to chr variables
  
```
Tidy data! and convert route~ from rows to column and stored their values as route numbers
  
```{r Tidy_data}
 
  NYC_df =
          NYC_df %>%
                pivot_longer(                                                    # Tidy data! 
                              route1:route11, 
                              names_to = "route_name", 
                              values_to = "route_num") %>%
    
                      relocate(route_name, route_num, .before = entrance_type)

```
  The NYC_Transit dataset provides information about New York city Transportation system. The data includes information about train lines, station name, station location, available routes, types of entrance, entry and vending information. The data also includes ADA compliance of the various stations.
  Some steps I have taken to clean this data include, cleaning variable names, selecting the variables I need, and converting select variables to a format that would be useful and easy to manipulate. I also tidied the data, converting the route names from numerous rows into a single column thereby improving the readability of my dataset. Finally, I moved a variable around to keep with the flow of data. The dimension of the resulting dataset is (`r dim(NYC_df)`). The resulting data is tidier than the initial data.

There are `r NYC_df %>% distinct(line, station_name) %>% nrow()` distinct stations in this dataset and `r NYC_df %>% distinct(line, station_name, ada)%>% filter(ada == TRUE) %>% nrow()` of these stations are ADA compliant. The proportion of station entrances that allow entrance without vending is `r NYC_df %>%distinct(line, station_name, entry, vending) %>% filter(entry == TRUE, vending == "NO") %>% nrow()`. There are `r NYC_df %>%distinct(line, station_name, route_name, route_num) %>% filter(route_num == "A") %>% nrow()` distinct stations that serve the A train. Of these, only `r NYC_df %>%distinct(line, station_name, route_name, route_num, ada) %>% filter(route_num == "A", ada == TRUE) %>% nrow()` of them are ADA compliant. 

# Problem 3

Read dataset pols-month.csv, clean the datasets, and use function 'separate' to split "mon" into year, month and day. Also, convert from character variable to numeric. 

```{r read_pols_month_df, message = FALSE}

path_4_data = "./Data/fivethirtyeight_datasets/pols-month.csv"

  pols_month_df =
  
                  read_csv(
                           path_4_data) %>%
    janitor::clean_names() %>%
      separate(col = mon, into = c("Year", "Month", "Day"), sep = "-") %>%
        mutate(Year = as.numeric(Year),
               Month = as.numeric(Month),
               Day = as.numeric(Day)) 
  

```
Replace Month number with Month name by creating a tibble helper and then perform a left_join *(Note to self: This Month_df variable differs from month_df in Trashwheel. This is uppercased the other is not)

```{r Month_modify}

Month_df =
    tibble(
            Month = 1:12,
            Month_name = month.name)

    pols1_month_df =
                    left_join(pols_month_df, Month_df, by = "Month") %>%
                      relocate(Month_name, .before = Day) 
          
```
